### 2022-04-23 chapters/chapter3.tex
switch conclusion to discussion
update

\chapter{Episodic Self-Imitation Learning with Hindsight}
\section{Introduction}
Reinforcement learning (RL) has has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the  agent. Reward shaping, however, often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robotic manipulation tasks, manually designing  reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations to the task are introduced. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals. The complicated dynamics of robot manipulation exacerbate the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require successfully executing multiple steps over a long horizon involves high dimensional control {that must also generalise to work across variations in the environment for each step}. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it is not able to learn at all. A common solution in the robotics community is to collect a sufficient quantity of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the achievable performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to enhance training, and this is particularly useful in sparse reward environments.

To alleviate the problems associated with having sparse rewards, there are two kinds of approaches: imitation learning and hindsight experience replay (HER). First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari Games. Whilst being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, {the present} experiments suggest that SIL struggles when rewards are sparse. Recently, {hindsight experience replay} has been proposed to solve such goal-conditional, sparse reward problems. The main idea of HER~\cite{andrychowicz2017hindsight} is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

In this paper, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm}: proximal policy optimization (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts entire current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises entire current collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the Open AI Fetch environment, this paper demonstrates that the proposed ESIL approach is effective in training agents which are required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption{Illustration of difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

The primary contribution of this paper is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by~\cite{plappert2018multi}. {The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-conditional settings than the standard self-imitation learning algorithm. Finally, this approach achieves, to our knowledge, the best results for four moderately complex robot control tasks in simulation.

% Related work
\section{Related Work}
% exploration
Imitation learning (IL) can be divided into two main categories: behavioural cloning and inverse reinforcement learning~\cite{hussein2017imitation}. Behavioural cloning involves the learning of behaviours from demonstrations~\cite{bojarski2016end,xu2017end,torabi2018behavioral}. Other extensions have an expert in the loop, such as DAgger~\cite{ross2011reduction}, or use an adversarial paradigm for the behavioural cloning method~\cite{ho2016generative,wang2017robust}. The inverse reinforcement learning estimates a reward model from expert trajectories~\cite{ng2000algorithms,abbeel2004apprenticeship,ziebart2008maximum}. Learning from demonstrations is powerful for complex robotic manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}. ~\cite{ho2016generative} propose generative adversarial imitation learning (GAIL), which employs generative adversarial training to match the distribution of state--action pairs of demonstrations. Compared with behavioural cloning, the GAIL framework shows strong improvements in continuous control tasks. In the work of~\cite{ding2019goal}, goalGAIL is proposed to speed up the training in goal-conditional environments; goalGAIL was also shown to be able to learn from demonstrations without action information. Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming}. Demonstrations are often collected by an expert policy or human actions. In contrast to these approaches, {episodic self-imitation learning (ESIL) does not need demonstrations.}

% self-learning and self-imitation learning
Self-imitation learning (SIL)~\cite{oh2018self} is used for exploiting past experiences for parametric policies. It~has a similar flavor to ~\cite{gangwani2018learning, pmlr-v97-wu19a}, in that the agent learns from imperfect demonstrations. During~training, past good experiences are stored in the replay buffer. When SIL starts, transitions are sampled from the replay buffer according to the advantage values. {In the work of~\cite{tang2020self}, generalised~SIL was proposed as an extension of SIL. It uses an $n$-bound $Q$-learning approach to generalise the original SIL technique, and shows robustness to a wide range of continuous control tasks. Generalised SIL can also be combined with both deterministic and stochastic RL algorithms. ~\cite{guo2019self} points out that using imitation learning with past good experience could lead to a sub-optimal policy. Instead of imitating past good trajectories, a trajectory-conditioned policy~\cite{guo2019self} is proposed to imitate trajectories in diverse directions, encouraging exploration in environments where exploration is otherwise difficult.} Unlike SIL, episodic self imitation learning (ESIL) applies HER to the current episodes to create ``imperfect'' demonstrations for imitation learning; this also requires introducing a trajectory-selection module to reject undesired samples from the hindsight experiences. In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling the sparse or delayed reward environments. {The present} experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.

% Hindsight
Recently, the technique known as hindsight learning was developed. Hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is an algorithm that can overcome the exploration problems in multi-goal environments, delivering sparse rewards. Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-conditional policies using hindsight experiences. However, the current implementation of HPG has only been evaluated for agents that need to perform discrete actions, and one drawback of hindsight policy gradient estimators is the computational cost because of the goal-oriented sampling. An extension of HER, called dynamic hindsight experience replay (DHER)~\cite{fang2018dher}, was proposed to deal with dynamic goals. ~\cite{liu2019hindsight} uses the GAIL framework~\cite{ho2016generative} to generate trajectories that are similar to hindsight experiences; it then applies imitation learning, using these trajectories. Competitive Experience Replay (CER) complements HER by introducing a competition between two agents for exploration~\cite{liu2018competitive}. ~\cite{zhao2018energy} point out that the hindsight trajectories which contain higher energy are more valuable during training, leading to a more efficient learning system. ~\cite{fang2019curriculum} proposed curriculum-guided HER, which~ incorporates curriculum learning in the work. During training, the agent focuses on the closest goals in the initial stage, then focuses on the expanding the diversity of goals. This approach accelerates training compared with other baseline methods. Unlike these works, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the start of training. Furthermore, ESIL can be applied to continuous control, making it more suitable for control problems that demand greater precision.

% Methodology
\section{Methodology}
\label{sec:method}
The proposed method combines PPO and episodic self-imitation learning to maximally use hindsight experiences for exploration to improve learning. {Recent advantages in episodic backward update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.

\subsection{Episodic Self-Imitation Learning}
The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ time steps; after $T$, a series of transitions $\tau_i =\{ (s_{t}|\langle g, g^{ac}_{t} \rangle, a_{t}, r_{t}, s_{t+1}|\langle g, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected. If at time step $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}|\langle g^\prime, g^{ac}_{t} \rangle, a_{t}, r_{t}^{\prime},s_{t+1}| \langle g^\prime, g^{ac}_{t+1}\rangle )\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimization (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as
\begin{equation}
  \centering
  \mathcal{L} = \alpha\cdot\mathcal{L}_{PPO} + \beta\cdot\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\cdot \mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\eta$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}A_{t}, \clip \left (\frac{\pi_{\theta}(a_{t}|s_{t}, g)}{\pi_{\theta_{old}}(a_{t}|s_{t}, g)}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $R_{t} - V_{\eta}(s_{t}, g)$. $V_{\eta}(s_{t}, g)$ is the state value at time step $t$ which is predicted by the critic network. $R_{t}$ is the return at time step $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is an squared error loss $\mathcal{L}_{value}=(V_{\eta}(s_{t}, g) - R_{t})^{2}$.}

For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $L_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning
\begin{equation}
  \beta=\frac{N_{ESIL}}{N_{Total}},
\end{equation}
where $N_{ESIL}$ is the number of samples used for self-imitation learning and $N_{Total}$ is the total number of collected samples.
The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as 
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\cdot\mathcal{F}_t\right],
  \label{eq:imitaion}
\end{equation}
where $\Tau^{'}$ indicates hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $R$, and the returns of corresponding hindsight experiences, $R^{\prime}$.

\subsection{Episodic Update with Hindsight}
Two important issues of ESIL are: (1) hindsight experiences are sub-optimal, and (2) the {detrimental} effect of updating networks with correlated trajectories. Although episodic self-imitation learning makes exploration more effective, hindsight experiences are not from experts and not ``perfect'' demonstrations.  With the training process continuing, if the agent is always learning these imperfect demonstrations, the policy will be stuck at the sub-optimal, or experience overfitting.

To prevent the agent learning from imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at time step $t$, the return $R_{t}$ can be calculated by $R_{t} = R(s_t, g) = r_{t} + \gamma \cdot r_{t+1} + \gamma^{2} \cdot r_{t+2} + ... + \gamma^{T-t-1} \cdot r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{R_{0}^{i}, R_{1}^{i}, R_{2}^{i}, ..., R_{T-1}^{i}\right\}$.

>>>>

\chapter{Episodic Self-Imitation Learning with Hindsight}
\label{ch:esil}
\section{Introduction}
% \td{In this chapter, we propose episodic self-imitation learning (ESIL), which aims to address }
%------------------------------ From Main Paper ----------------------------
Reinforcement learning (RL) has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the agent. However, reward shaping often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robot manipulation tasks, the manual design of reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations are introduced in the task. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals. The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require the successful execution of multiple steps over a long horizon involves high-dimensional control that must also be generalised to work across variations in the environment for each step. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all. A common solution in the \td{DRL} community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to improve training, which is particularly useful in sparse reward environments.

There are two kinds of approaches to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance the exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari games. While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse. Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems. The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

In this chapter, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm:} proximal policy optimisation (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts the entire set of current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises the entire currently collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

The primary contribution of this chapter is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by Plappert \textit{et al.}~\cite{plappert2018multi} -- ``\textit{How to combine HER with state-of-the-art on-policy RL algorithms like
PPO?}". {The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-oriented settings than the standard self-imitation learning algorithm. 
% Finally, to our knowledge, this approach achieves the best results for four moderately complex robot control tasks in simulation. 
Finally, this approach outperforms other on-policy RL algorithms in all simulated robot manipulation tasks.

% Related work
\section{Related Work}
% exploration
Imitation learning (IL) can be divided into two main categories: behavioural cloning and inverse reinforcement learning~\cite{hussein2017imitation}. Behavioural cloning involves the learning of behaviours from demonstrations~\cite{bojarski2016end,xu2017end,torabi2018behavioral}. Other extensions have an expert in the loop, such as DAgger~\cite{ross2011reduction}, or use an adversarial paradigm for the behavioural cloning method~\cite{ho2016generative,wang2017robust}. The inverse reinforcement learning estimates a reward model from expert trajectories~\cite{ng2000algorithms,abbeel2004apprenticeship,ziebart2008maximum}. Learning from demonstrations is powerful for complex robot manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}. Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming},  nevertheless demonstrations are often collected by an expert policy or human actions. In contrast to these approaches, {episodic self-imitation learning (ESIL) does not need demonstrations.}

% self-learning and self-imitation learning
Self-imitation learning (SIL)~\cite{oh2018self} can be used for exploiting past experiences for parametric policies. It~has a similar flavor to ~\cite{gangwani2018learning, pmlr-v97-wu19a}, in that the agent learns from imperfect demonstrations. During~training, past good experiences are stored in the replay buffer. When SIL starts, transitions are sampled from the replay buffer according to the advantage values. {In the work of~\cite{tang2020self}, generalised~SIL was proposed as an extension of SIL. It uses an $n$-bound $Q$-learning approach to generalise the original SIL technique, and shows robustness to a wide range of continuous control tasks. Generalised SIL can also be combined with both deterministic and stochastic RL algorithms. Guo \textit{et al.}~\cite{guo2019self} points out that using imitation learning with past successful experiences could lead to a sub-optimal policy. Instead of imitating past good trajectories, a trajectory-conditioned policy~\cite{guo2019self} is proposed to imitate trajectories in diverse directions, encouraging exploration in environments where exploration is otherwise difficult.} Unlike SIL, episodic self imitation learning (ESIL) applies HER to the current episodes to create ``imperfect'' demonstrations for imitation learning; this also requires introducing a trajectory-selection module to reject undesired samples from the hindsight experiences. In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling in the sparse or delayed reward environments. The present experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.

% Hindsight
Recently, the technique known as hindsight learning has been developed. Hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is an algorithm that can overcome the exploration problems in multi-goal environments, delivering sparse rewards. Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-oriented policies using hindsight experiences with on-policy RL algorithms. However, the current implementation of HPG has only been evaluated for agents that need to perform discrete actions, and one drawback of hindsight policy gradient estimators is the computational cost because of the goal-oriented sampling. Unlike HPG, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the beginning of training. Furthermore, ESIL can be applied to continuous control, making it more suitable for control problems that demand greater precision.

% Methodology
\section{Methodology}
\label{sec:method}
The proposed method combines PPO and episodic self-imitation learning to maximally use hindsight experiences for exploration to improve learning. {Recent advantages in episodic update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.

\subsection{Episodic Self-Imitation Learning}
The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ timesteps; after $T$ timesteps, a series of transitions $\tau_i =\{ (s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected. If at timestep $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimisation (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as:
\begin{equation}
  \centering
  \mathcal{L} = \alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as:
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\psi$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as:
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}A_{t}, \clip \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $G_{t} - V(s_{t}, g;\psi)$. $V(s_{t}, g;\psi)$ is the state value at timestep $t$ which is predicted by the critic network. $G_{t}$ is the return at timestep $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is a squared error loss $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t}, g;\psi) - G_{t})^{2}]$.}

For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $\mathcal{L}_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning:
\begin{equation}
  \beta=\frac{N_{ESIL}}{N_{Total}},
\end{equation}
where $N_{ESIL}$ is the number of samples used for self-imitation learning and $N_{Total}$ is the total number of collected samples.
The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as: 
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi\left(a_{t}|s_{t}, g^\prime;\theta\right)\mathcal{F}_t\right],
  \label{eq:imitaion}
\end{equation}
where $\Tau^{'}$ denotes hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $G$, and the returns of corresponding hindsight experiences, $G^{\prime}$.

\subsection{Episodic Update with Hindsight}
Two crucial issues of ESIL are: 1) hindsight experiences are sub-optimal, and 2) the {detrimental} effect of updating networks with correlated trajectories. 
Although episodic self-imitation learning makes exploration more effective, hindsight experiences are not from experts and not ``perfect'' demonstrations.  With the training process continuing, if the agent is always learning these imperfect demonstrations, the policy will be stuck at the sub-optimal, or experience overfitting.

To prevent the agent from learning imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at each timestep $t$, the return $G_{t}$ can be calculated by $G_{t} = G(s_t, g) = r_{t} + \gamma r_{t+1} + \gamma^{2} r_{t+2} + ... + \gamma^{T-t-1} r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$.

For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated. Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$. During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be \td{assumed to be of lower importance} and ignored.
Then, Equation~\eqref{eq:imitaion} can be rewritten as:
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
  \label{eq:final_loss}
\end{equation}
where $\mathcal{F}\left(s_{t}, g, g^{\prime}\right)$ is the trajectory selection module. The selection function can be expressed as
\begin{equation}
  \centering
  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[G(s_{t}, g^\prime) > G(s_{t}, g)\right],
\end{equation}
here, {$\mathds{1}(\cdot)$ is the unit step function}. Consider the OpenAI Fetch environment as an example. For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$. {The desired goal is modified} to construct a new hindsight trajectory, and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$. Then,  {$G$ and $G^{\prime}$ can be calculated~separately.}
\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
  \caption{A simplified illustration of trajectory selection. Blue trajectories indicate original experiences. Orange trajectories indicate hindsight experiences. Solid trajectories in the hindsight experiences are selected by the trajectory selection module of ESIL with new ``imagined'' goals.}
  \label{fig:hs_princple}
\end{figure}

From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals. The full description of ESIL can be found in Algorithm~\ref{alg}.
% the algorithm
\begin{algorithm}[t!]
  \caption{Proximal Policy Optimisation (PPO) with Episodic Self-Imitation Learning (ESIL)}
  \label{alg}
  \begin{algorithmic}[1]
    \REQUIRE an actor network $\pi(s, g;\theta)$, a critic network $V(s, g;\psi)$, the maximum steps $T$ of an episode, a reward function $\mathcal{R}$, number of iterations $N$, number of episodes for each iteration $M$, weight coefficient of critic loss term $c$, weight coefficient of PPO term $\alpha$, adaptive coefficient $\beta$  
    %\STATE Initialize the parameters $\theta$ and $\psi$ of both actor and critic networks
    \FOR{$\mathrm{iteration} = 1, 2, \cdots N$}
    \STATE $\Tau \leftarrow \varnothing$, $\Tau^{\prime} \leftarrow \varnothing$ 
    \FOR{$\mathrm{episode} = 1, 2, ..., M$}
    \STATE $\tau \leftarrow \varnothing$
    \FOR{$t = 0, 1, \cdots, T-1$}
    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g;\theta)$
    \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g^{ac}_{t+1}\right)$ in $\tau$
    \ENDFOR
